---
title: "Stat 184 Final Project"
author: "Ashley Reed"
subtitle: "Chicago Crime: 2001 to Present"
output: html_notebook
---

In this project, I will be analyzing data collected by the Chicago Police Department that was posted for public access on the Chicago Data Portal. This data set contained 6.75 million cases providing information on 22 different variables. Each case represents a reported crime. 

The variables include:

* ID (Unique ID numbers for each reported crime)
* Case number (Unique combinations of numbers and letters used by the Chicago Police Department to identify cases)
* Date
* Block
* IUCR (Illinois Uniform Crime Reporting code)
* Primary Type (general description of IUCR code)
* Description (detailed description of IUCR code)
* Location Description
* Arrest (true or false)
* Domestic (true of false)
* Beat
* District
* Ward
* Community Area
* FBI Code
* X-coordinate (location)
* Y-coordinate (location)
* Year
* Updated On (date of last update)
* Latitude
* Longitude 
* Location

### Front Matter

```{r}
rm(list = ls())

library(rvest)
library(tidyverse)
library(tidyr)
library(mosaic)
library(lubridate)
library(DataComputing)
library(readr)
library(lubridate)
library(knitr)

```

### Reading in the Data

```{r, results = "hide"}

ChicagoCrimeOriginal <-
read_csv("Crimes_-_2001_to_present.csv")

# Reading in data from the Chicago Data Portal to create a new data frame in R --------------------------------------------

``` 

```{r}

ChicagoCrimeOriginal

```

### Cleaning the Data

In order to clean up the data to make it tidy and ready for analysis, I renamed the variables to fit standard data conventions in R programming, changed variable formats to fit what they contain (ex. dates, logical values, etc.), and removed the X or XX included in addresses under the block variable.

```{r}

ChicagoCrime <-
ChicagoCrimeOriginal %>%
  rename(id = ID, 
         case_number = `Case Number`, 
         date = Date, 
         block = Block, 
         iucr = IUCR, 
         primary_type = `Primary Type`,
         description = Description,
         location_description = `Location Description`,
         arrest = Arrest,
         domestic = Domestic,
         beat = Beat,
         district = District,
         ward = Ward,
         community_area = `Community Area`,
         fbi_code = `FBI Code`,
         x_coordinate = `X Coordinate`,
         y_coordinate = `Y Coordinate`,
         year = Year,
         updated = `Updated On`,
         lat = Latitude,
         long = Longitude,
         location = Location) %>%
  mutate(date = lubridate::mdy_hms(date),
         arrest = as.logical(arrest),
         domestic = as.logical(domestic),
         beat = as.integer(beat),
         district = as.integer(district),
         block = gsub("X|XX","", block))

# Cleaning Chicago Crime data to fit standard conventions and simplify further work  -------------------------------------

ChicagoCrime 

```

### Analysis

Now that I have cleaned up the data, I will use data wrangling to analyze the data and come to some meaningful conclusions. I'm curious to find out more about crime rates (specifically in terms of arrests) over the years, so I will create a plot of years vs. number of crimes, and plot crimes resulting in arrest and crimes not resulting in arrest seperately.

```{r}

ChicagoCrimeYears <-
  ChicagoCrime %>%
  group_by(year, arrest) %>%
  summarise(count = n()) 

ChicagoCrimeYears %>%
  ggplot(aes(x = year, y = count, color = arrest)) +
  geom_point() +
  geom_line() 

```

The resulting plot comes out with some very interesting results.  While I had expected that crime rates would show an increase as the years progressed, this shows the exact opposite. The number of crimes committed each year in Chicago is at a high from around 2001-2005, and then proceeds to decrease relatively consistently, except for a few isolated peaks, until it reaches an ultimate minimum value in 2018.  

In addition, it seems that many more crimes take place in Chicago that lead to no arrest that crimes resulting in an arrest. I had assumed the opposite would be true when creating this plot. In certain years, there were almost triple the number of crimes with an arrest that did not result in arrest.  

Overall, this plot could be seen as relatively good news; from the trends shown here, we could reasonable hypothesize that crimes will continue to decrease in Chicago in the coming years.

------------------------------------------------------------

Next, I would like to observe the difference in number of total crimes committed over the years in each Chicago District. 


```{r}
 
ChicagoCrimeAssault <-
ChicagoCrime %>%
  select(district, arrest) %>%
  group_by(district, arrest) %>%
  summarise(count = n()) 

ChicagoCrimeAssault %>%
  filter(arrest == "TRUE") %>%
  arrange(desc(count)) %>%
  head(5)

ChicagoCrimeAssault %>%
  filter(arrest == "FALSE") %>%
  arrange(desc(count)) %>%
  head(5)

ChicagoCrimeAssault %>%
  ggplot(aes(x = district, y = count)) +
  geom_col(aes(fill = arrest), position = "dodge") +
  scale_x_continuous(breaks = c(1, 3, 5, 7, 9, 11, 13, 15, 17, 19, 21, 23, 25), limits = c(1,25)) 

```

month of year, top 4 types, facet by arrest and types

```{r}

pattern <- "\\d{4}\\-(\\d{2})"

ChicagoCrimeMonth <-
ChicagoCrime %>%
  filter(grepl(pattern, date)) %>%
  tidyr::extract(date, into = "month", regex = pattern) %>%
  mutate(month = as.integer(month))

ChicagoCrime %>%
  group_by(primary_type) %>%
  summarise(count = n()) %>%
  filter(rank(desc(count)) <= 4)

ChicagoMonthGlyphReady <-
ChicagoCrimeMonth %>%
  select(primary_type, arrest, month) %>%
  filter(primary_type %in% c("THEFT", "BATTERY", "NARCOTICS", "CRIMINAL DAMAGE")) %>%
  group_by(primary_type, arrest, month) %>%
  summarise(count = n()) 

```


```{r}
  
ChicagoMonthGlyphReady %>%
  ggplot(aes(x = month, y = count)) +
  geom_col(aes(fill = arrest), position = "dodge") +
  scale_x_continuous(breaks = c(1:12)) +
  facet_wrap(primary_type ~ .)

```
